---
title: "Antibiotic Resistance based on ARIBA output"
output: html_notebook
---

Ariba output was generated using the [MegaRes](https://megares.meglab.org) database for comparison.

```{r}
library(staphopia)
library(dplyr)
library(ggplot2)
```

switch to dev version.  ignore when the database is moved to new version.
```{r}
USE_DEV = TRUE
```

Get table of all public samples - takes a minute or so.  There are 43972 samples.
```{r}
ps <- get_public_samples()
```

Create a mega-table of all ARIBA output for every sample.
```{r}
start.time <- Sys.time()
ariba_tab <- get_resistance_results(ps$sample_id)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```
This is the raw [ARIBA report file output](https://github.com/sanger-pathogens/ariba/wiki/Task:-run)

Becasue this file takes a while to load, I created a feather datafile for saving and fast reloading

```{r}
library(feather)
write_feather(x = ariba_tab, path = "./ariba_tab.feather")
```

properties of this dataframe

```{r}
nrow(ariba_tab)
select(ariba_tab,sample_id) %>% unique() %>% nrow()
```


### Tetracycline resistance

```{r}
tet <- filter(ariba_tab,class == 10) %>%
  group_by(sample_id,cluster_name) %>%
  summarise(tot_reads = sum(as.integer(reads)))
  
```

exploratory dotplot of reads versus type.  Shows that some genes have very high variance - prob a feature of plasmid carriage

```{r}
bp <- ggplot(tet,aes(x=cluster_name,y=tot_reads)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_y_log10()
bp
```

Get a list of all strains with a least one hit, any coverage.  Can use this chunk to later play around with sensitivity by tweaking parameters.

```{r}
tet_hits <- select(tet,sample_id) %>%
  unique() 
```

Merge with public samples and label

```{r}
tet_res <- ps %>%
  select(sample_id,st) %>%
  mutate(tetracycline = ifelse(sample_id %in% tet_hits$sample_id,"resistant","sensitive"))
```

Get top sequence types

```{r}
tst <- get_top_sequence_types()
```

Make summary of tet resistance

```{r}
tet_summary <- group_by(tet_res,st,tetracyline) %>%
  summarise(count = sum(tetracycline))
```


## Beta Lactam resistance


```{r}
bla <- filter(ariba_tab,class == 13) %>%
  group_by(sample_id,mechanism) %>%
  summarise(tot_reads = sum(as.integer(reads)))
  
```

Visualization


```{r}
bp2 <- ggplot(bla,aes(x=mechanism,y=tot_reads)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_y_log10()
bp2
```

```{r}
betalactamases_hits <- filter(bla,grepl("betalactam",mechanism)) %>%
  select(sample_id) %>%
  unique() 
```

```{r}
pbp_hits <- filter(bla,grepl("Penicillin",mechanism)) %>%
  select(sample_id) %>%
  unique() 
```

### mecA specifically

```{r}
mecA_hits <- filter(ariba_tab,grepl("MEC[A|-]",cluster_name)) %>%
  select(sample_id) %>%
  unique() 
```



```{r}
sessionInfo()
```

