---
title: "Public Sequencing Rank By Year"
output:
  pdf_document: default
---

This is a quick example on generating plots of the publicly available *Staphylococcus aureus* sequencing data by year.

### Import Packages

```{r}
library(staphopia)
library(ggplot2)
library(reshape2)
USE_DEV = TRUE
```

### Get Our Data
We'll use the **get_rank_by_year()** function to do exactly that, retrieve submission counts by year.

```{r}
results <- get_rank_by_year()
results
```

In the table above, there are seven columns:

1. year: The year in which an experiment was made public in ENA/SRA
2. bronze: The number of bronze ranked samples for a given year
3. silver: The number of silver ranked samples for a given year
4. gold: The number of gold ranked samples for a given year
5. count: The number of submissions for a given year
6. overall_bronze: The sum of bronze ranked samples of the previous years
7. overall_silver: The sum of silver ranked samples of the previous years
8. overall_gold: The sum of gold ranked samples of the previous years
9. overall: The sum of each of the previous years

### Plotting Our Data
We'll use *ggplot2* to visualize our data. We'll look at the number of Bronze, Silver and Gold ranked samples of 

#### Published vs Unpublished By Year
For our final plot, we'll look at the number of samples that were referenced in a publication along side those that weren't. We'll need to melt the data in order to plot our groups.
```{r fig.width=12, fig.asp=0.5}
melted <- melt(results, id=c('year'),
               measure.vars = c('bronze', 'silver', 'gold'))
melted$title <- ifelse(melted$variable == 'gold', 'Gold', 
                       ifelse(melted$variable == 'silver', 'Silver', 'Bronze'))
melted$rank <- ifelse(melted$variable == 'gold', 3, 
                      ifelse(melted$variable == 'silver', 2, 1))

title <- substitute(paste("Sequencing ranks (Bronze = ", b, ", Silver = ", s,
                          ", Gold = ", g, ") of publicly available ",
                          italic('S. aureus')," samples between ", min_year,
                          " and ", max_year, "."), list(
    b=format(max(results$overall_bronze), big.mark=',', scientific=FALSE),
    s=format(max(results$overall_silver), big.mark=',', scientific=FALSE),
    g=format(max(results$overall_gold), big.mark=',', scientific=FALSE),
    min_year=min(results$year),
    max_year=max(results$year)
))
p <- ggplot(data=melted, aes(x=year, y=value, fill=title, group=rank, label=title)) +
    xlab("Year") +
    ylab("Sequencing Rank Per Year") +
    ggtitle(title) +
    geom_bar(stat='identity', position='dodge') +
    geom_text(aes(label=value), vjust = -0.5, position = position_dodge(.9)) +
    scale_fill_manual(values=c("#CD7F32", "#D4AF37", "#C0C0C0")) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"),
          legend.title = element_blank())
    

p
```



```{r}
sessionInfo()
```

