---
title: "Public Submissions By Year"
output: html_notebook
---

# Public Submissions By Year
This is a quick example on generating plots of the publicly available *Staphylococcus aureus* sequencing data by year.

### Import Packages

```{r}
library(staphopia)
library(ggplot2)
```

### Get Our Data
We'll use the **get_submission_by_year()** function to do exactly that, retrieve submission counts by year.

```{r}
results <- get_submission_by_year()
results
```

In the table above, there are three columns:

1. year: The year in which an experiment was made public in ENA/SRA
2. count: The number of submissions for a given year
3. overall: The sum of each of the previous years

### Plotting Our Data
We'll use *ggplot2* to make two simple plots to visualize our data.

#### Submissions Per Year
Here we are just going to look at the per year submissions.

```{r}
p <- ggplot(data=results, aes(x=year, y=count)) +
    xlab("Year") +
    ylab("Total Submitted Samples") +
    geom_bar(stat='identity') +
    geom_text(aes(label=count), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw()
p
```


#### Overall Submissions
Changing gears a litte, here we are going to look at the overall growth of *S. aureus* sequencing data over the years.

```{r}
p <- ggplot(data=results, aes(x=year, y=overall)) +
    xlab("Year") +
    ylab("Overall Submitted Samples") +
    geom_bar(stat='identity') +
    geom_text(aes(label=overall), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw()
p
```

