---
title: "Public Submissions By Year"
output: 
  html_notebook: 
    fig_height: 6
    fig_width: 12
---

This is a quick example on generating plots of the publicly available *Staphylococcus aureus* sequencing data by year.

### Import Packages

```{r}
library(staphopia)
library(ggplot2)
library(reshape2)
```

### Get Our Data
We'll use the **get_submission_by_year()** function to do exactly that, retrieve submission counts by year.

```{r}
results <- get_submission_by_year()
results
```

In the table above, there are seven columns:

1. year: The year in which an experiment was made public in ENA/SRA
2. published: The number of submissions associated with a publication for a given year
3. unpublished: The number of submissions **not** associated with a publication for a given year
4. count: The number of submissions for a given year
5. overall_published:The sum of published samples of the previous years
6. overall_unpublished: The sum of unpublished samples of the previous years
7. overall: The sum of each of the previous years

### Plotting Our Data
We'll use *ggplot2* to make two simple plots to visualize our data.

#### Submissions Per Year
Here we are just going to look at the per year submissions.

```{r fig.width=12, fig.asp=0.4}
title <- substitute(paste("Per year (2010-present) total of publicly available sequenced ", italic('S. aureus')," samples (N = ", x,")."), list(x=format(max(results$overall), big.mark=',', scientific=FALSE)))
p <- ggplot(data=results, aes(x=year, y=count)) +
    xlab("Year") +
    ylab("Total Submitted Samples Per Year") +
    ggtitle(title) +
    geom_bar(stat='identity') +
    geom_text(aes(label=count), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```


#### Overall Submissions
Changing gears a litte, here we are going to look at the overall growth of *S. aureus* sequencing data over the years.

```{r fig.width=12, fig.asp=0.4}
title <- substitute(paste("Cumulative total of publicly available sequenced ", italic('S. aureus')," samples (N = ", x,") by year (2010-present)."), list(x=format(max(results$overall), big.mark=',', scientific=FALSE)))
p <- ggplot(data=results, aes(x=year, y=overall)) +
    xlab("Year") +
    ylab("Cumulative Submitted Samples") +
    ggtitle(title) +
    geom_bar(stat='identity') +
    geom_text(aes(label=overall), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Published vs Unpublished By Year
For our final plot, we'll look at the number of samples that were referenced in a publication along side those that weren't. We'll need to melt the data in order to plot our groups.
```{r fig.width=12, fig.asp=0.5}
melted <- melt(results, id=c('year'), measure.vars = c('published', 'unpublished'))
melted$title <- ifelse(melted$variable == 'published', 'Published', 'Unpublished')

title <- substitute(paste("Published (N = ", p, ") and unpubished (N = ", u, ") publicly available ", italic('S. aureus')," samples between ", min_year, " and ", max_year, "."), list(
    p=format(max(results$overall_published), big.mark=',', scientific=FALSE),
    u=format(max(results$overall_unpublished), big.mark=',', scientific=FALSE),
    min_year=min(results$year),
    max_year=max(results$year)
))
p <- ggplot(data=melted, aes(x=year, y=value, fill=title)) +
    xlab("Year") +
    ylab("Published and Unpublished Samples Per Year") +
    ggtitle(title) +
    geom_bar(stat='identity', position='dodge') +
    geom_text(aes(label=value), vjust = -0.5, position = position_dodge(.9)) +
    scale_fill_manual(values=c("#2ca25f", "#5ab4ac")) +
    scale_x_continuous(breaks = round(seq(min(results$year), max(results$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"),
          legend.title = element_blank())
    

p
```
