---
title: "Top Sequence Types"
output: html_notebook
---

This is a quick example on generating plots of top multi-locus sequence types (MLST) represented in the database.

### Import Packages

```{r}
library(staphopia)
library(ggplot2)
```

### Get Our Data
In order to retrieve the top sequence types we'll need to use the **get_top_sequence_types()** function.

```{r}
results <- get_top_sequence_types()
results
```

In the table above, there are three columns:

1. sequence_type: The MLST representation
2. count: Total number of samples with said seqeunce type

#### Adjusting Total Sequence Types Returned
By default **get_top_sequence_types()** only returns the top 10 sequence types. However, by giving **get_top_sequence_types()** an integer value, you can adjust how many sequence types are returned.

##### Example: Only the Top Sequence Type
You would use *get_top_sequence_types(**1**)*, notice the **1** value.
```{r}
results <- get_top_sequence_types(1)
results
```

##### Example: Only the Top 5 Sequence Types
You would use *get_top_sequence_types(**5**)*, notice the **5** value.
```{r}
results <- get_top_sequence_types(5)
results
```

##### Example: Top 'X' Sequence Types
You would use *get_top_sequence_types(**X**)*, where *X* is some integer value.
```{r}
X <- 3
results <- get_top_sequence_types(X)
results
```

### Plot Our Data
Here we'll generate a simple plot to visualize the top 10 sequence types represented in Staphopia's database.

```{r}
results <- get_top_sequence_types()
p <- ggplot(data=results, aes(x=reorder(sequence_type, -count), y=count)) +
    xlab("Sequence Type") +
    ylab("Total Represented Samples") +
    geom_bar(stat="identity") +
    geom_text(aes(label=count), vjust = -0.5) + 
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```
